---
title: "R Manual"
output: 
  html_notebook:
    theme: paper
    toc: yes
    toc_float: yes
---
This manual is an R cheatsheet for Psychology students. 
It explains how to complete many tasks psychologists commonly do.
Click through the table of contents to quickly find the task you need.

# How to install and load pacakges {.tabset}

## Haven
Haven contains functions that lets R import SPSS, SAS, Stata, and other foreign files.

Documentation can be viewed [here](https://haven.tidyverse.org)
```{r}
# Check if haven is already installed and if it is, load it.
if (!require(haven)){
  # If it's not intalled, then tell R to install it.
  install.packages("haven", dependencies = TRUE)
  # Once it's installed, tell R to load it.
  library(haven)
}
```

# Import an SPSS file
Use the read_sav() function from the Haven package to import an SPSS file into R.
```{r}
dataset <- read_sav("ExampleDataset")
```

# View a dataset
Use the View() function to look at your dataset like you would in SPSS or Excel
```{r}
View(dataset)
```

