---
title: "R Manual"
output: 
  html_notebook:
    theme: paper
    toc: yes
    toc_float: yes
---
This manual is an R cheatsheet for Psychology students. 
It explains how to complete many tasks psychologists commonly do.
Click through the table of contents to quickly find the task you need.

# How to install and load pacakges {.tabset}

<iframe width="560" height="315" src="https://www.youtube.com/embed/hndMg7ldBJk" frameborder="0" allow="autoplay; encrypted-media" allowfullscreen></iframe>

## Haven
Haven contains functions that lets R import SPSS, SAS, Stata, and other foreign files.

Documentation can be viewed [here](https://haven.tidyverse.org)
```{r}
# Check if haven is already installed and if it is, load it.
if (!require(haven)){
  # If it's not intalled, then tell R to install it.
  install.packages("haven", dependencies = TRUE)
  # Once it's installed, tell R to load it.
  library(haven)
}
```

## Tidyverse
Tidyverse contains many functions useful to cleaning, tidying, and manipulating data

Documentation can be viewed [here](https://www.tidyverse.org)

```{r}
if (!require(tidyverse)){
  install.packages("tidyverse", dependencies = TRUE)
  library(tidyverse)
}
```

## summarytools

Summarytools provide useful functions for summarizing and visualizing data

Documentation can be viewed [here](https://www.rdocumentation.org/packages/summarytools/versions/0.8.5)
```{r}
if (!require(summarytools)){
  install.packages("summarytools", dependencies = TRUE)
  require(summarytools)
}
```
## psych

Summarytools provide useful functions for summarizing and visualizing data

Documentation can be viewed [here](https://www.rdocumentation.org/packages/summarytools/versions/0.8.5)
```{r}
if (!require(psych)){
  install.packages("psych", dependencies = TRUE)
  require(psych)
}
```


# Import an SPSS file

<iframe width="560" height="315" src="https://www.youtube.com/embed/8zpHc9U8HIY" frameborder="0" allow="autoplay; encrypted-media" allowfullscreen></iframe>

Use the read_sav() function from the Haven package to import an SPSS file into R.

Documentation can be viewed [here](https://www.rdocumentation.org/packages/haven/versions/1.1.2/topics/read_spss)
```{r}
dataset <- read_sav("https://osf.io/98mt6/download")
```

# Clean dataset {.tabset}

<iframe width="560" height="315" src="https://www.youtube.com/embed/gFALxkGfPQQ" frameborder="0" allow="autoplay; encrypted-media" allowfullscreen></iframe>

## View dataset
Use the View() function to look at your dataset like you would in SPSS or Excel

Documentation can be viewed [here](https://www.rdocumentation.org/packages/utils/versions/3.5.1/topics/View)

```{r}
View(dataset)
```
## Remove columns
Use the Select() function to de-select a column.

Put '-' before the column name to remove a column (you can list more than one column).

Just type the column name to keep only that column (you can list more than one column).

Documentation can be viewed [here](https://www.rdocumentation.org/packages/dplyr/versions/0.7.6/topics/select)

```{r}
dataset %>%
  select(-IPAddress) -> dataset
```

## Remove practice runs
Use the slice() funciton to remove rows.

Put the range of rows you'd like to remove with minus signs in front.

Documentation can be viewed [here](https://www.rdocumentation.org/packages/dplyr/versions/0.7.6/topics/slice)

```{r}
dataset %>% 
  slice(-1:-3) -> dataset
```

# Summarize variables {.tabset}

## Use dfSummary function from summarytools package to summarize and visualize data

<iframe width="560" height="315" src="https://www.youtube.com/embed/L8IwS7shipg" frameborder="0" allow="autoplay; encrypted-media" allowfullscreen></iframe>

Documentation can be viewed [here](https://www.rdocumentation.org/packages/summarytools/versions/0.8.5/topics/dfSummary)

```{r}
#First select the variables you'd like to summarize
dataset %>%
  select (CoinFlip, FFM_5, Potter3) -> exampleDF

#Then print them with this command
print(dfSummary(exampleDF, graph.magnif = .75), method = 'render')
```
## Use the describe() function of the psych package to summarize data

<iframe width="560" height="315" src="https://www.youtube.com/embed/K83_tYXqFVo" frameborder="0" allow="autoplay; encrypted-media" allowfullscreen></iframe>

Documentation can be viewed [here](https://www.rdocumentation.org/packages/psych/versions/1.8.4/topics/describe)

```{r}
#First select the variables you'd like to summarize
dataset %>%
  select (CoinFlip, FFM_5, Potter3) %>%
  describe()
```
# Create Composite Variable {.tabset}

## Psych {.tabset}

Documentation can be viewed [here](https://www.rdocumentation.org/packages/psych/versions/1.8.4/topics/scoreItems)

### Extraversion
```{r}
#create dataframe with only relevant variables to work with
Extraversion <- data.frame (dataset$FFM_1, dataset$FFM_6, dataset$FFM_11, dataset$FFM_16, dataset$FFM_21, dataset$FFM_26, dataset$FFM_31, dataset$FFM_36)
#create list of 'keys'. The  numbers just refer to the order of the question in the data.frame() you just made. The most important thing is to mark the questions that should be reversed scored with a '-'. 
Extraversion.keys <- make.keys(Extraversion, list(Extraversion=c(1,-2,3,4,-5,6,-7,8)))
#score the scale
Extraversion.scales <- scoreItems (Extraversion.keys, Extraversion)
#save the scores
Extraversion.scores <- Extraversion.scales$scores
#save the scores back in 'dataset'
dataset$Extraversion <- Extraversion.scores[,]
#print the cronbach alpha
Extraversion.scales$alpha
```
### Agreeableness
```{r}
#create dataframe with only relevant variables to work with
Agreeableness <- data.frame (dataset$FFM_2, dataset$FFM_7, dataset$FFM_12, dataset$FFM_17, dataset$FFM_22, dataset$FFM_27, dataset$FFM_32, dataset$FFM_37, dataset$FFM_42)
Agreeableness.keys <- make.keys(Agreeableness, list(Agreeableness=c(-1,2,-3,4,5,-6,7,-8,9)))
Agreeableness.scales <- scoreItems (Agreeableness.keys, Agreeableness)
Agreeableness.scores <- Agreeableness.scales$scores
dataset$Agreeableness <- Agreeableness.scores[,]
Agreeableness.scales$alpha
```
### Conscientiousness
```{r}
#create dataframe with only relevant variables to work with
Conscientiousness <- data.frame (dataset$FFM_3, dataset$FFM_8, dataset$FFM_13, dataset$FFM_18, dataset$FFM_23, dataset$FFM_28, dataset$FFM_33, dataset$FFM_38, dataset$FFM_43)
my.keys <- make.keys(Conscientiousness, list(Conscientiousness=c(1,-2,3,-4,-5,6,7,8,-9)))
my.scales <- scoreItems (my.keys, Conscientiousness)
my.scores <- my.scales$scores
dataset$Conscientiousness <- my.scores[,]
my.scales$alpha
```
### Neuroticism
```{r}
#create dataframe with only relevant variables to work with
Neuroticism <- data.frame (dataset$FFM_4, dataset$FFM_9, dataset$FFM_14, dataset$FFM_19, dataset$FFM_24, dataset$FFM_29, dataset$FFM_34, dataset$FFM_39)
my.keys <- make.keys(Neuroticism, list(Neuroticism=c(1,-2,3,4,-5,6,-7,8)))
my.scales <- scoreItems (my.keys, Neuroticism)
my.scores <- my.scales$scores
dataset$Neuroticism <- my.scores[,]
my.scales$alpha
```
### Openness
```{r}
#create dataframe with only relevant variables to work with
Openness <- data.frame (dataset$FFM_5, dataset$FFM_10, dataset$FFM_15, dataset$FFM_20, dataset$FFM_25, dataset$FFM_30, dataset$FFM_35, dataset$FFM_40, dataset$FFM_41, dataset$FFM_44)
my.keys <- make.keys(Openness, list(Openness=c(1,2,3,4,5,6,-7,8,-9,10)))
my.scales <- scoreItems (my.keys, Openness)
my.scores <- my.scales$scores
dataset$Openness <- my.scores[,]
my.scales$alpha
```
## Tidyverse

Documentation can be viewed [here](https://www.rdocumentation.org/packages/dplyr/versions/0.7.6/topics/rowwise) and [here](https://www.rdocumentation.org/packages/plyr/versions/1.8.4/topics/mutate)

```{r}
dataset %>%
  rowwise() %>%
  mutate(extraversionTidy = mean(c(FFM_1, 6-FFM_6, FFM_11, FFM_16, 6-FFM_21, FFM_26, 6-FFM_31, FFM_36))) -> dataset
```