---
title: "R Manual"
output: 
  html_notebook:
    theme: paper
    toc: yes
    toc_float: yes
---
This manual is an R cheatsheet for Psychology students. 
It explains how to complete many tasks psychologists commonly do.
Click through the table of contents to quickly find the task you need.

# How to install and load pacakges {.tabset}

## Haven
Haven contains functions that lets R import SPSS, SAS, Stata, and other foreign files.

Documentation can be viewed [here](https://haven.tidyverse.org)
```{r}
# Check if haven is already installed and if it is, load it.
if (!require(haven)){
  # If it's not intalled, then tell R to install it.
  install.packages("haven", dependencies = TRUE)
  # Once it's installed, tell R to load it.
  library(haven)
}
```

## Tidyverse
Tidyverse contains many functions useful to cleaning, tidying, and manipulating data

Dcoumentation can be viewed [here](https://www.tidyverse.org)

```{r}
if (!require(tidyverse)){
  install.packages("tidyverse", dependencies = TRUE)
  library(tidyverse)
}
```

# Import an SPSS file
Use the read_sav() function from the Haven package to import an SPSS file into R.
```{r}
dataset <- read_sav("https://osf.io/98mt6/download")
```

# Clean dataset {.tabset}

## View dataset
Use the View() function to look at your dataset like you would in SPSS or Excel
```{r}
View(dataset)
```
## Remove columns
Use the Select() function to de-select a column.

Put '-' before the column name to remove a column (you can list more than one column).

Just type the column name to keep only that column (you can list more than one column).
```{r}
dataset %>%
  select(-IPAddress) -> dataset
```

## Remove practice runs
```{r}
dataset %>% 
  slice(-1:-3) -> dataset
```

